---
title: 'Data Management Plan'
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: 
    reference_docx: "sci_paper_urblogo.dotx"
urlcolor: blue
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(readr)
library(readxl)
library(qwraps2)
library(gridExtra)
library(kableExtra)
library(bibliometrix)
```

## 0 Administrative data

**Version of DMP**: v1.0

**Project title:** *Bibliometric Analysis of European Research on
Digital Divide: An Exploration of the Corporate Landscape*

**Start and end of project:** 01.11.2022 - 15.02.2023

```{r nrows, include=FALSE}
load(here("Script", "Environments", "Normalize_env.RData"))


MM <- M


```

**Data summary:** This bibliometric aims to examine the state of the art
of European research in the field of the digital divide by combining
data from the Web of Science, Scopus, and Dimensions bibliographic
platforms. Additionally, this work seeks to explore the corporate
digital divide. The limit of the search comprises authors with European
affiliations within the business, management, economics, technology, and
computer science disciplines. After processing, merging, and cleaning, a
total of `r nrow(MM)` unique documents were incorporated into the final
data set, including articles, book chapters, conferences, and
proceedings. The results will be obtained by operating the R programming
language using the bibliometrix package and the biblioshiny application.

## 1. Data Description

This research project will use secondary data collected by conducting a
specific search on the digital divide using the Web of Science, Scopus,
and Dimensions platforms. The three platforms have different graphic
user interfaces that delimit the search and the supported formats in
which bibliographical data is downloaded. The content of bibliographical
data varies between text, numeric, and integers data types, and the
formats that generate each platform will be readable using the R
programming language

## 1.1 Data collection

Before collecting the data, access to the graphical user interfaces of
each platform was granted through university affiliation. During this
stage, the search criteria, the queries, and the formats downloaded are
the following:

```{r wos, include=FALSE}

load(here("Script", "Environments", "Biblio_data_bases.RData"))

WW <- W

```

#### Web of Science

This platform allows one to choose the formats and the fields. While
exporting the data, a *plain text file* was chosen. The custom selection
used all the Web of Science core collection fields for the record
content.

-   **Searched Fields:** Keywords and Title
-   **Searched Text:** "digital divide\*" OR "digital inequalit\*" OR
    "digital gap\*"
-   **Document Types:** Articles, proceedings book chapters, review
    articles and early access
-   **Web of Science Categories:** Computer science and technology,
    management, business and economics.
-   **Region:** Countries in Europe
-   **Time Frame:** `r min(WW$PY, na.rm = T)` -
    `r max(WW$PY, na.rm = T)`
-   **Date of Search:** 10.10.2022
-   **Total Documents:** `r nrow(WW)`
-   **Query:** [Go to Web of Science Query
    Link](https://www.webofscience.com/wos/woscc/summary/a10a1d1b-bcfa-4494-9ed0-75228e2f40e1-6839a671/date-descending/1)
-   **Downloaded Data Type:** .txt
-   **Download data and Size:** [eu_wos_1\_500.txt 2.5
    MB](https://www.dropbox.com/s/qlkh336108iy0nj/eu_wos_1_500.txt?dl=0)
    and [eu_wos_501_1032.txt 2.8
    MB](https://www.dropbox.com/s/ppqw2az5mcik5j4/eu_wos_501_1032.txt?dl=0)

### Scopus

This database also allows one to choose the formats and the necessary
fields to conduct a bibliometric analysis. In this case, all the
categories of citation information, bibliographical information,
abstracts, and keywords.

```{r scopus, include=FALSE}

load(here("Script", "Environments", "Biblio_data_bases.RData"))


SS <- S
```

-   **Searched Fields:** Keywords and Title
-   **Searched Text:** "digital divide\*" OR "digital inequalit\*" OR
    "digital gap\*"
-   **Document Types:** Articles, proceedings, book chapters, review
    articles, and early access articles.
-   **Scopus Categories:** Computer science and technology, management,
    business and economics.
-   **Region:** Countries in Europe.
-   **Time Frame:** `r min(SS$PY, na.rm = T)` -
    `r max(SS$PY, na.rm = T)`
-   **Date of Search:** 05.10.2022
-   **Total Documents:** `r nrow(SS)`
-   **Query:** [Vizualize
    query](https://www.dropbox.com/s/m53l0v0z1i21kc4/scopus_query.rtf?dl=0)
-   **Downloaded Data Type:** .csv
-   **Download data and Size:** [eu_scopus_1\_1786.csv 16.9
    MB](https://www.dropbox.com/s/fywg0lnvq243zve/eu_scopus_1_1786.csv?dl=0)

### Dimensions


This database has a less developed graphic user interface. Even though
it lets one choose the format, it does not let one choose the fields.
However, the search can be customized using the Application Programming
Interface API.

```{r dim, include=FALSE}
load(here("Script", "Environments", "Biblio_data_bases.RData"))

DD <- D
```

-   **Searched Fields:** Title and Abstract.
-   **Searched Text:** "digital divide\*" OR "digital inequalit\*" OR
    "digital gap\*"
-   **Document Types:** Articles, proceedings, and book chapters.
-   **Dimensions Categories:** (38) Economics, (35) Commerce,
    Management, Tourism and Services, and (46) Information and Computing
    Sciences.
-   **Region:** Countries in Europe.
-   **Time Frame:** `r min(DD$PY, na.rm = T)` -
    `r max(DD$PY, na.rm = T)`
-   **Date of Search:** 14.10.2022
-   **Total Documents:** `r nrow(DD)`
-   **Query:** Not available for graphic user interface.
-   **Downloaded Data Type:** .csv
-   **Download data and Size:** [dim_1\_1467.csv 8.3
    MB](https://www.dropbox.com/s/35wfrt4kp9ujoct/dim_1_1467.csv?dl=0)
    and [dim_1468_3108.csv 3.8
    MB](https://www.dropbox.com/s/te4ykjln8l6gihu/dim_1468_3108.csv?dl=0)

## 1.2 Data Processing

```{r col_names, include=FALSE}
ww_col <- ncol(WW)
ss_col <- ncol(SS)
dd_col <- ncol(DD)
mm_col <- ncol(MM)

```

The R programming language environment will be used during the data
processing stage. After downloading, the raw files (.txt and .csv
formats) from each platform will be converted into a bibliographic
database format using the bibliometrix package. The volume of the
generated data in this repository will be estimated between 0.7 to 1 GB.

After converting the downloaded datasets from the three platforms into
bibliographic data frames, there were differences in column length and
names, meaning the three platforms used different codebooks for naming
the variables. For example, the [Web of Science
codebook](https://www.dropbox.com/scl/fi/2kw4wge5y7prwfn8i6mzb/col_names_wos.xlsx?dl=0&rlkey=panwseiwrktiut611n6cpu04m)
contained `r ww_col` variables, the [Scopus
codebook](https://www.dropbox.com/scl/fi/4ochrul4cu5d0v4xsgrdb/col_names_scopus.xlsx?dl=0&rlkey=milgv2adjegonkom45yfunxeb)
`r ss_col` variables, and the [Dimensions
codebook](https://www.dropbox.com/scl/fi/0t01j093rwtsqxozna9dj/col_names_dim.xlsx?dl=0&rlkey=o87k0k0ddy3cy2ga0ugyy0u4k)
`r dd_col` variables. To solve this problem, first, we followed the [Web
of Science Core Collection
codebook](https://support.clarivate.com/ScientificandAcademicResearch/s/article/Web-of-Science-Core-Collection-List-of-field-tags-in-output?language=en_US)
to homogenize the variable's names. Second, the [bibliometrix
manual](https://www.bibliometrix.org/vignettes/Data-Importing-and-Converting.html)
suggests selecting the main variables to conduct a bibliometric
analysis. As a result, the final bibliographic [data
frame](https://www.dropbox.com/s/w6y29syjhqs2e88/M_EU.csv?dl=0) contains
`r mm_col` variables ([see
codebook](https://www.dropbox.com/scl/fi/ls7et6b55epo5g5t5ecmk/colnames_M.xlsx?dl=0&rlkey=j0we18gy90mnnlfawnp00jlxq)).

This
[script](https://github.com/luchocastillo84/bibliometrix_dd_europe/blob/master/Script/Biblio_data_bases.R)
contains the R-Studio workflow used to process, clean, and merge the
bibliographic data frames from the three sources.

## 2. Documentation and Data Quality

The metadata creation will follow the Data Documentation Initiative DDI
standards, including a codebook that describes the data and the variable
types. To guarantee replicability both in the data collection and
output, the reuse of data, and high-quality results, the following
measures will be taken: pretests, data validation, peer review of data,
repeat measurements, and the use of software and digital tools. This
repository,
[bibliometric_digital_divide_Europe](https://github.com/luchocastillo84/bibliometrix_dd_europe),
will follow the R-Studio workflow using the bibliometrix package and the
biblioshiny application for data processing and analysis. An ordered
structure of the folders with documented code will also be managed using
the version control offered by GitHub to track changes and manipulate
data responsibly.

## 3. Storage and technical archiving

This research project does not contain any sensible data; thus,
throughout the project duration, data will be stored on the
collaborative workspace R-Studio, GitHub, and Dropbox.

## 4. Responsibilities and resources

This bibliometric study is the initial part of a research agenda
proposed by the principal researcher as academic work, which will be
presented at the end of his doctoral studies. The principal investigator
will collect, analyze, present, and write the study results in a
scientific article format. In addition, this project will be oriented
and guided by the supervisor and the co-supervisor.

### Principal Investigator and author of the DMP:

Luis Carlos Castillo-Tellez

University of Urbino

[l.castillotellez\@campus.uniurb.it](mailto:l.castillotellez@campus.uniurb.it){.email}

[ORCID](https://orcid.org/my-orcid?orcid=0000-0003-2041-4596)

### Supervisor:

Francesco Vidoli

University of Urbino

[francesco.vidoli\@uniurb.it](mailto:francesco.vidoli@uniurb.it){.email}

[ORCID](https://orcid.org/0000-0003-4125-9337)

### Co-Supervisor:

Michael Rochlitz

University of Bremen

[michael.rochlitz\@uni-bremen.de](mailto:michael.rochlitz@uni-bremen.de){.email}

[ORCID](https://orcid.org/0000-0001-8652-8874)
