---
title: 'Bibliometric Analysis of European Research on Digital Divide: An Exploration of the Corporate Landscape'
author: "Luis Carlos Castillo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  beamer_presentation:
    keep_tex: true
    citation_package: natbib
    includes:
      in_header: mystyle2.tex
  slidy_presentation: default
  powerpoint_presentation: default
institute: |
  | University of Urbino
  | Ph.D. Program in Global Studies
editor_options:
  markdown:
    wrap: 72
bibliography: references.bib
link-citations: yes
linkcolor: myblue
natbiboptions: round
header-includes:
  - \usepackage{etoolbox}
  - \AtBeginEnvironment{thebibliography}{\scriptsize}
biblio-style: "apalike"
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F )
kableExtra.auto_format = FALSE
knitr.table.format = "latex"
knitr::opts_knit$set(kable.force.latex = TRUE)
```

```{r Loading packages, include=FALSE}
library(tidyverse)
library(devtools)
library(here)
library(bibliometrix)
library(stringr)
library(readr)
library(visdat)
library(stringr)
library(dplyr)
library(readxl)
library(janitor)
library(ggplot2)
library(gridExtra)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(rmarkdown)
library(tidytext)
library(knitr)
library(qwraps2)
library(dplyr)
library(tables)
library(bibtex)
library(igraph)
library(RColorBrewer)
library(Matrix)
library(flextable)
library(ggpubr)
library(cowplot)
library(ggeasy)
library(Hmisc)
library(gridGraphics)
library(gridBase)

```

```{r Functions, include=FALSE}

source("coc_mat_func.R")
source("netplot_func.R")




```

# Content I

1.  Digital Divide Overview

2.  Motivation

3.  Objectives and Research Questions

4.  Data

5.  Bibliometric Analysis

6.  Performance Analysis

    1.  Publications Vs Citations
    2.  Authors
    3.  Articles
    4.  Journals
    5.  Affiliations/ Universities
    6.  Countries

# Content II

7.  Science Mapping

    1.  Citations Analysis
    2.  Similarity Measures
        1.  Co-citations Analysis
        2.  Bibliographic Coupling
       

8.  Conclusions

# 1. The Digital Divide Overview I

-   The digital divide is also known as the digital gap, inequalities,
    or disparities.
-   The interaction with other existing gaps such as income, education,
    gender, age, and regional, among others [@ragnedda2017].
-   The evolution of the concept has pointed out the phenomenon's
    complexity and the effects on the different layers of society and
    the economy [@vandijk2003; @ragnedda2017; @shakina2021].

# 1. The Digital Divide Overview II

-   Waves of Research
    -   **The first wave:** Physical access to technology -\> possession
        of computers and access to the internet [@norris2001;
        @james2002; @castells2003].

    -   **The second wave:** Usage of digital technologies and skills
        [@hargittai2002b; @vandijk2005b; @vandijk2006c;
        @vandeursen2011c].

    -   **The third wave:** The disparities in tangible outcomes arise from different forms of access to and usage of digital technologies, emphasizing the ability to benefit from these technologies in a data-driven market to enhance personal and professional aspects (Ragnedda, 2017; van Deursen, 2015).

# 1. The digital divide overview VI

## The corporate landscape

-   Digital revolution -\> different aspects of daily activities -\> how
    we conduct business.
-   Disparity in digital capabilities and resources among businesses
    [@shakina2021].
-   The corporate digital divide is a topic that remains under-explored
    [@pejicbach2013; @shakina2021].
-   Understanding and addressing the divide, policymakers, and
    businesses owners can target their efforts to ensure inclusive
    digital transformation.

# 2. Motivation

-   Investigating the transformative effects of digital technologies on society and economy, while highlighting both opportunities and challenges.

-   Aligning with the Digital Europe program's vision by devising strategies to bridge the digital divide effectively.

-   Diversifying bibliometric research by extending its application beyond health sciences, computer science, and technology to understand the digital divide.

-   Harnessing the power of comprehensive data from three leading academic platforms to generate insightful and actionable findings on the digital divide.

# 3. Objectives and Research Questions I

## Objectives

-   Understand the intellectual structure within the domain of the
    digital divide.

-   Examine the intellectual interactions,
    and thematic relationships of European research components.

-   Explore the corporate digital divide among the collected corpus and
    identify trends and patterns within the literature.

# 3. Objectives and Research Questions II

## Research Questions

-   Q1: How have the main trends, focus shifts, and key themes in European research on the digital divide evolved over time, and how do they reflect the current state of knowledge in this field?

-   Q2: What are the intellectual interactions and thematic relationships among European research components on the digital divide, and how do they contribute to the identification of core subtopics and literature clusters?

-   Q3: How are European studies addressing the corporate digital divide, and which unexplored topics within this domain warrant further examination?


# 4. Data I

```{r Total docs, include=F}
MM <- read_csv(here("Data", "Processed", "M_EU.csv"))
rownames(MM) <- MM$SRDI # using SR as row names
class(MM) <- c("bibliometrixDB", "data.frame")
```

-   Specific search within titles and author keywords on the "*digital
    divide*" merging data from the Web of Science, Scopus, and
    Dimensions platforms.
-   Search criteria: "digital divide\*" OR "digital inequalit\*" OR
    "digital gap\*"
-   The sample includes articles, book chapters, conferences, and
    proceeding papers.
-   Authors with European affiliations within the **business,
    management, economics, technology, and computer science**
    disciplines were included

# 4. Data II

-   After conducting a thorough data cleaning, a total of `r nrow(MM)`
    unique documents from `r min(MM$PY)` to `r max(MM$PY)` were
    incorporated.

-   Number of Documents by Database

    -   WoS:`r sum(MM$PF == "WOS")`
    -   Scopus: `r sum(MM$PF == "SCO")`
    -   Dimensions: `r sum(MM$PF == "DIM")`

-   To track the evolution of the digital divide literature, the data
    was divided into three periods: 2000-2007, 2008-2015, and 2016-2022.

-   The R programming language environment(bibliometrix, igraph
    packages) was used to carry out the analysis.

# 5. Methodology I

## Bibliometric Analysis

Following @donthu2021, @Aria2017, @ellegaard2015 and @Bornmann2015
bibliometric analysis:

-   Is a methodology that applies quantitative techniques to
    bibliographic data and plays a vital role in evaluating research
    output.
-   This technique allows researchers to uncover emerging trends
    identifying knowledge gaps in specific domains and analyze a significant quantity of documents .
-   It offers three types of analysis: performance analysis, science
    mapping, and network analysis.

# 5. Methodology II

\vspace{0.5cm}

## Distribution of Document Types Across Time Periods


```{r doc types, include=TRUE}
load(here("Script", "Environments",  "Tables&Graphs.RData"))

 ggplot(g_rates_2, aes(x = Period, y = as.numeric(Value), fill = Details)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = Value, 
                group = Details), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5,
            size = 3) +
  theme_minimal() +
  labs(title = "Document Types by Periods",
    x = "Periods",
    y = "Documents",
    fill = "Doc Types") +
  scale_fill_manual(values = c("#012a4a", "#0466c8", "#468faf","#7d8597")) +
  theme(legend.position = "top") +
   theme(plot.title = element_text(hjust = 0.5)) +
  theme(text = element_text(size = 14)) 
```

# 6. Performance Analysis I

Is a descriptive interpretation of research constituents.

## 6.1. Publications vs Citations

\vspace{0.5cm}

```{r Cite vs Prod}
cite <- ggplot(tot_cite_pub, aes(x= PY, y= log(avg_cite_pery))) + 
  geom_line() + 
  geom_vline(xintercept = c(2008, 2015), linetype = "dotted",color = "blue") + 
  labs(title = "B. Relative Change in the Average Citations per Year",
       y = "Log Avg Citations", x ="Years") + 
  theme(text = element_text(size = 14)) 

prod <- ggplot(tot_cite_pub, aes(x= PY, y= totd)) + 
  geom_line() + 
  geom_vline(xintercept = c(2008, 2015),linetype = "dotted",color = "blue") + 
  labs(title = "A. Total Publications per Year" ,
       y = "Total Publications", x ="Years") + 
  theme(text = element_text(size = 14)) 
```

```{r Panel cite vs prod, fig.height=7, fig.width=13}

grid.arrange(prod, cite, nrow= 1)
```

# 6. Performance Analysis II

\vspace{0.5cm}




# 6. Performance Analysis IX

## 6.6. Countries Collaboration Network

```{r Network CO, include=T}
load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(40993)
col_CO <- invisible(net_CO <- netPlot(CC_CO_mat, 
                                      n = 40, Title = "Country Collaboration Network",
            type = "fruchterman", size.cex=TRUE, size=20, 
            remove.multiple=FALSE, labelsize=0.9,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain", remove.isolates = T,
            legend.position = "bottomright", legend.keywords = coutries_clus))


```

# 7. Science Mapping I

Is a set of techniques and tools used to visualize and analyze the structure, relationships, and patterns within a scientific field or discipline.


## 7.1. Citation Analysis

```{r Most cited refs, include=T}
most_Lcr <- read_csv(here("Data", "Processed", "most_Lcr.csv"), show_col_types = F)

kbl(most_Lcr, "latex",  caption = "Most Cited References 2000-2022") %>% 
   column_spec(2, latex_column_spec = "p{9cm}") %>% 
  # column_spec(4, latex_column_spec = "p{5cm}") %>% 
kable_styling(bootstrap_options = "striped", full_width = F, font_size = 5,
              latex_options = "hold_position") %>%
  row_spec(0, bold=T) %>% add_footnote(c("Source: Author's elaboration", "TC: Times Cited"), notation ="number") 
```

# 7. Science Mapping II

## 7.2. Similarity measures

Quantify similarity, connections and relationships among academic entities.

Following @kammerer2021

-   **Knowledge base:** cluster of academic publications in a research
    field that are considered fundamental to the development and
    understanding of the field.

-   **Research front:** cluster of academic publications that refers to
    emerging active areas of research considering themselves with a
    similar unsolved research problem.

```{=tex}
\begin{center}
\includegraphics[width=0.6\textwidth]{pic_1.png}
\end{center}
```

# 7. Science Mapping III

\vspace{0.5cm}

### 7.2.1. Co-citations Analyisis


```{r Co_cite_P1,include=T}

load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(30100)
net=netPlot(CC1_mat, n= 35, Title = "Co-Citation Network,  2000 - 2007", # Louvain
                type = "fruchterman", size.cex=TRUE, size=20, 
                remove.multiple= FALSE, labelsize=0.9,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain", curved = F,
            legend.position = "topright", legend.keywords = keywords1)

```

# 7. Science Mapping IV

```{r Co_cite_P2,include=T}

load(here("Script", "Environments",  "Net4Mat.RData"))


set.seed(1030034)
net=netPlot(CC2_mat, n = 30, Title = "Co-Citation Network, 2008 - 2015", 
            type = "fruchterman", size.cex=TRUE, size=20, 
            remove.multiple=FALSE, labelsize=0.9,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain",
            legend.position = "topright", legend.keywords = keywords2)

```

# 7. Science Mapping V

```{r Co_cite_P3,include=T}

load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(1003)
net=netPlot(CC3_mat, n = 30, Title = "Co-Citation Network,  2016 - 2022", 
            type = "fruchterman", size.cex=TRUE, size=20, label.color = F,
            remove.multiple=FALSE, labelsize=0.7,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain", legend.position = "bottomright", 
            legend.keywords = keywords3)

```

# 7. Science Mapping VI

## Co-citation Networks Summary

This three networks highlighted:

-   **Evolution of research focus:** From internet access and infrastructure to 
the various facets of the digital divide.

-   **Key authors and publications:** These form the knowledge foundations that shapes the discourse, broaden the comprehension of the literature, and guide future research.

-   **Emerging trends and themes:** These networks have assisted in uncovering growing themes that highlight research directions that require further examination.


# 7. Science Mapping VII

## 7.2.2. Bibliographic Coupling Analysis

```{r Bib_coup_P1,include=T}

load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(9001)
net=netPlot(BC1_mat, n = 30, Title = "Bibliographic Coupling Network, 2000-2007", 
            type = "fruchterman", size.cex=TRUE, size=20, 
            remove.multiple=FALSE, labelsize=0.7,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain", remove.isolates = T, legend.position = "topleft", 
            legend.keywords = keywords4)
```

# 7. Science Mapping VIII

```{r Bib_coup_P2,include=T}

load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(90001)
net=netPlot(BC2_mat, n = 30, Title = "Bibliographic Coupling Network, 2008-2015", 
            type = "fruchterman", size.cex=TRUE, size=15, 
            remove.multiple=FALSE, labelsize=0.7,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain", legend.position = "bottomleft", 
            legend.keywords = keywords5)

```

# 7. Science Mapping IX

```{r Bib_coup_P3,include=T}

load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(1001)
net=netPlot(BC3_mat, n = 30, Title = "Bibliographic Coupling Network, 2016-2022", 
            type = "fruchterman", size.cex=TRUE, size=15, 
            remove.multiple=FALSE, labelsize=0.7,edgesize = 3, edges.min=1, label = T,
            cluster = "louvain", legend.position = "bottomright", 
            legend.keywords = keywords6)

```

# 7. Science Mapping X

## Bibliographic coupling Networks Summary

This three networks highlighted:

-   **Evolution of digital divide research:** Themes move from access and socio-demographic 
factors to a more nuance understanding of skills, usage and emerging technologies.

-   **Emergence of key authors:** Represent the influential authors that contributed in the emergence of active areas of research in the development of the digital divide.

-   **Growing complexities and specialization:** The diverse research themes in the networks reflect the expanding scope and depth of digital divide research.


# 7. Science Mapping XI

## 7.2.2. Co-word Analysis

```{r co_word_P1,include=F}

load(here("Script", "Environments",  "Net4Mat.RData"))

# set.seed(6900012)
# net=netPlot(WCo1_mat, n = 40, Title = "Co-Word Network 2000-2007", 
#             type = "fruchterman", size.cex=TRUE, size=15, 
#             remove.multiple=FALSE, labelsize=0.7,edgesize = 4, edges.min=1, label = T,
#             cluster = "louvain")
set.seed(6902012)
net=netPlot(WCo_mat, n = 50, Title = "Co-Word Network 2000-2022", 
            type = "fruchterman", size.cex=TRUE, size=15, 
            remove.multiple=FALSE, labelsize=0.7,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain")


```



```{r co_word_P2,include=F}
# 7. Science Mapping XI
load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(6900013)
net=netPlot(WCo2_mat, n = 30, Title = "Co-Word Network 2008-2015", 
            type = "fruchterman", size.cex=TRUE, size=15, 
            remove.multiple=FALSE, labelsize=0.7,edgesize = 3, edges.min=5, label = T,
            cluster = "louvain")

```



```{r co_word_P3,include=F}
# 7. Science Mapping XII
load(here("Script", "Environments",  "Net4Mat.RData"))

set.seed(6900014)
net=netPlot(WCo3_mat, n = 30, Title = "Co-Word Network 2016-2022", 
            type = "fruchterman", size.cex=TRUE, size=15, 
            remove.multiple=FALSE, labelsize=0.7,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain")

```

# Conclusions

-  We have seen the main trends and focus shifts: from access, infrastructure, and
socio-economic factors, to skills, usage and other facets of the digital divide, this highlights the complexity and multidimensionality of the digital divide.

-   The networks showcase collaboration among prominent authors that consistently contribute in the field. The thematic relationships show interconnections of various aspects of the digital divide.

-   European studies have not extensively addressed the corporate digital divide, leaving room for further examination. The corporate digital divide might be incorporated into other literature streams, such as digital transformation and technology adoption.


# Contact information 

\vspace{2cm}

Luis Carlos Castillo-Tellez

\vspace{0.1cm}

Guest Researcher

PhD candidate in Global Studies

University of Urbino - Italy

email: l.castillotellez@campus.uniurb.it

email: luiscarl@uni-bremen.de



# References {.allowframebreaks}

::: {#refs}
:::
