---
title: 'Bibliometric Analysis of European Research on Digital Divide: An Exploration of the Corporate Landscape'
author: "Luis Carlos Castillo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  beamer_presentation:
    includes:
      in_header: mystyle1.tex
  slidy_presentation: default
  powerpoint_presentation: default
institute: |
  | University of Urbino
  | Ph.D. Program in Global Studies
editor_options:
  markdown:
    wrap: 72
bibliography: references.bib
header-includes:
  - \AtBeginEnvironment{CSLReferences}{\tiny}
biblio-style: apalike
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F )
kableExtra.auto_format = FALSE
knitr.table.format = "latex"
knitr::opts_knit$set(kable.force.latex = TRUE)
```

```{r Loading packages, include=FALSE}
library(tidyverse)
library(devtools)
library(here)
library(bibliometrix)
library(stringr)
library(readr)
library(visdat)
library(stringr)
library(dplyr)
library(readxl)
library(janitor)
library(ggplot2)
library(gridExtra)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(rmarkdown)
library(tidytext)
library(knitr)
library(qwraps2)
library(dplyr)
library(tables)
library(bibtex)
library(igraph)
library(RColorBrewer)
library(Matrix)
library(flextable)
library(ggpubr)
library(cowplot)
library(ggeasy)
library(Hmisc)
library(gridGraphics)
library(gridBase)

```

```{r Functions, include=FALSE}

source("coc_mat_func.R")
source("netplot_func.R")




```

# Content I

1.  Motivation

2.  Objectives

3.  Digital Divide Overview

4.  Data

5.  Bibliometric Analysis

6.  Performance Analysis

    1.  Publications Vs Citations
    2.  Authors
    3.  Articles
    4.  Journals
    5.  Affiliations/ Universities
    6.  Countries

# Content II

7.  Science Mapping

    1.  Citations Analysis
    2.  Co-citations Analysis
    3.  Bibliographic Coupling
    4.  Co-word Analysis

# 1. Motivation

-   The ubiquitous use of digital technologies has transformed many
    aspects of society and the economy, bringing potential benefits and
    drawbacks.

-   The Digital Europe program and the agenda to bridge the digital
    divide.

-   The few studies using the bibliometric method focus mainly on health
    sciences, computer science, and technology.

-   Data availability from three platforms containing references and
    citations from academic journals has stimulated this research
    initiative.

# 2. Objectives and Research Questions I

## Objectives

-   Understand the intellectual structure within the domain of the
    digital divide.

-   Examine European research components' intellectual interactions,
    structural connections, and thematic relationships.

-   Explore the corporate digital divide among the collected corpus and
    identify trends and patterns within the literature.

# 2. Objectives and Research Questions II

## Research Questions

-   What have been the main trends and shifts in the focus of European
    research on the digital divide over time?

-   Which are the most prominent authors and publications in the field?

-   How do the intellectual interactions, structural connections, and
    thematic relationships among European research components on the
    digital divide shape the current state of knowledge in this field?

-   What are the key themes and subtopics related to the digital divide
    in European research, and how are they grouped or clustered in the
    literature?

-   How is the corporate digital divide addressed in European research
    on the digital divide, and what areas have been largely unexplored
    in this field?

# 3. The Digital Divide Overview I

-   The digital divide is also known as the digital gap, inequalities,
    or disparities.
-   The interaction with other existing gaps such as income, education,
    gender, generational, and regional, among others [@ragnedda2017].
-   The evolution of the concept has pointed out the phenomenon's
    complexity and the effects on the different layers of society and
    the economy [@vandijk2003; @ragnedda2017; @shakina2021].

# 3. The Digital Divide Overview II

-   Waves of Research
    -   **The first wave:** Physical access to technology -\> possession
        of computers and access to the internet [@vandijk2003].

    -   **The second wave:** Usage of digital technologies and skills
        [@vandijk2006].

    -   The divide is not exclusively a matter of access or possession
        of technology. It encloses the ability to effectively search,
        access, and evaluate information using digital technologies.

# 3. The Digital Divide Overview III

-   Waves of Research
    -   **The third wave:** expands upon the inequalities previously
        identified in the 1st and 2nd wave.

    -   Focuses on the benefits previously gained from access, skills,
        and usage of digital technologies.

    -   Emphasize on the ability to benefit from digital technologies in
        a data-driven market to improve personal and professional
        aspects [@ragnedda2017].

    -   van Deursen et al. (2015) highlight the disparities from the
        tangible outcomes gained from different forms of access and
        usage of digital technologies.

# 3. The digital divide overview VI

## The corporate landscape

-   Digital revolution -\> different aspects of daily activities -\> how
    we conduct business.
-   The corporate digital divide is the disparity in digital
    capabilities and resources between businesses that can effectively
    use DT and those that struggle to access and use it [@shakina2021].
-   The corporate digital divide occurs at the regional, industry-type,
    and business-size levels [@pejicbach2013]. Usage gap [@pliskin2006].

# 3. The digital divide overview VI

## The corporate landscape: research in Europe

-   Out of 1641 articles retrieved, a second search within this sample
    found 32 articles addressing the corporate digital divide.
-   Period 2000-2007: refers to the digital gap among firms sizes,
    especially in the access to - ICT and the implementation of
    e-business.

# 3. The digital divide overview VII

## The corporate landscape: research in Europe

-   Period 2008-2015: the focus shifted to analyzing firm readiness for
    digital technologies and skills. It also identified the factors that
    contribute to the digital divide among businesses.
-   Period 2000-2022: emphasize digital skills and the advantages of DT.
    It also shows evidence from different countries.

# 4. Data I

```{r Total docs, include=F}
MM <- read_csv(here("Data", "Processed", "M_EU.csv"))
rownames(MM) <- MM$SRDI # using SR as row names
class(MM) <- c("bibliometrixDB", "data.frame")
```

-   Specific search within titles and author keywords on the "*digital
    divide*" merging data from the Web of Science, Scopus, and
    Dimensions platforms.
-   Search criteria: "digital divide\*" OR "digital inequalit\*" OR
    "digital gap\*"
-   The sample includes articles, book chapters, conferences, and
    proceedings.
-   Authors with European affiliations within the business, management,
    economics, technology, and computer science disciplines were
    included

# 4. Data II

-   After conducting a thorough data cleaning, a total of `r nrow(MM)`
    unique documents from `r min(MM$PY)` to `r max(MM$PY)` were
    incorporated.

-   To track the evolution of the digital divide literature, the data
    was divided into three periods: 2000-2007, 2008-2015, and 2016-2022.

-   Document distribution

    -   WoS:`r sum(MM$PF == "WOS")`
    -   Scopus: `r sum(MM$PF == "SCO")`
    -   Dimensions: `r sum(MM$PF == "DIM")`

-   The R programming language environment was used to carry out the
    analysis.

# 5. Bibliometric Analysis I

Following @donthu2021, @Aria2017, @ellegaard2015 and @Bornmann2015
bibliometric analysis:

-   Is a methodology that applies quantitative techniques to
    bibliographic data and plays a vital role in evaluating research
    output.
-   This technique allows researchers to uncover emerging trends
    identifying knowledge gaps in specific domains.
-   Is useful when analyzing a significant quantity of documents.
-   It offers three types of analysis: performance analysis, science
    mapping, and network analysis.

# 5. Bibliometric Analysis II

## Descriptive Analysis

```{r Descriptive, echo=FALSE, include=T}
desc_a <- read_csv(here("Data", "Processed", "desc_analysis.csv"))


# set_flextable_defaults(
#   border.color = "gray")
# 
# flextable(head(desc_a)) %>% 
#   bold(part = "header") %>% 
#   add_footer_lines("Descriptive Analysis") %>% 
#   theme_vanilla()

kable(desc_a,"latex") %>% 
  kable_styling(latex_options="scale_down") %>% 
  row_spec(0, bold=T)
  
   

```

# 6. Performance Analysis I

## 6.1. Publications vs Citations

\vspace{0.5cm}

```{r Cite vs Prod}
cite <- MM %>% group_by(PY) %>% 
  summarise(totc = sum(TC)) %>%  
  ggplot(aes(x= PY, y= totc)) + geom_line() + 
  geom_vline(xintercept = c(2007, 2014),
             linetype = "dotted", color = "blue") + 
  labs( y = "Citationbs", x ="Years")  + ggtitle( "Total Citations per Year") +
  theme(plot.title = element_text(hjust = 0.5)) +  
  theme(text = element_text(size = 20))

prod <- MM %>% group_by(PY) %>% 
  summarise(totd = n()) %>%  
  ggplot(aes(x= PY, y= totd)) + geom_line() + 
  geom_vline(xintercept = c(2008, 2015),linetype = "dotted", color = "blue") + 
  labs( y = "Publications", x ="Years")  + ggtitle( "Total Publications per Year") +
  theme(plot.title = element_text(hjust = 0.5)) +  
  theme(text = element_text(size = 20))
```

```{r Panel cite vs prod, fig.height=7, fig.width=13}

grid.arrange(prod, cite, nrow= 1)
```

# 6. Performance Analysis I

\vspace{0.5cm}

## 6.2. Evolving patterns in authors' scientific production over time.

```{r Prod of AU, include=TRUE}
load(here("Script", "Environments",  "Tables&Graphs.RData"))
p
```

# 6. Performance Analysis

## 6.2. Trends in authors' citations over time

```{r Influence of AU chart, include=TRUE}
load(here("Script", "Environments",  "Tables&Graphs.RData"))

c_bind
```

# 6. Performance Analysis

## 6.2. Authors II

```{r H-index AU chart, include=TRUE}
load(here("Script", "Environments",  "Tables&Graphs.RData"))
H_au
```

# 6. Performance Analysis

## 6.3. Articles I

```{r Influence of AU table , echo=FALSE, warning=FALSE}
influ_arti_bind <- read_csv(here("Data", "Processed", "influ_arti_bind.csv"), show_col_types = F)

kbl(influ_arti_bind, "latex", caption = "Most Cited Articles") %>% 
   column_spec(2, latex_column_spec = "p{9cm}") %>% 
  # column_spec(4, latex_column_spec = "p{5cm}") %>% 
kable_styling(bootstrap_options = "striped", full_width = F, font_size = 5) %>%
  row_spec(0, bold=T) %>%
  pack_rows("Period 1: 2000- 2007", 1, 5, latex_align = "c") %>%
  pack_rows("Period 2: 2008- 2015", 6, 10, latex_align = "c") %>% 
  pack_rows("Period 3: 2016- 2022", 11, 15, latex_align = "c") %>%
  add_footnote(c("Source: Author's elaboration", "TC: Times Cited"), notation ="number")
```

# 6. Performance Analysis

## 6.3. Articles II

```{r Most cited refs, include=T}
most_Lcr <- read_csv(here("Data", "Processed", "most_Lcr.csv"), show_col_types = F)

kbl(most_Lcr, "latex",  caption = "Most Cited References") %>% 
   column_spec(2, latex_column_spec = "p{9cm}") %>% 
  # column_spec(4, latex_column_spec = "p{5cm}") %>% 
kable_styling(bootstrap_options = "striped", full_width = F, font_size = 5,
              latex_options = "hold_position") %>%
  row_spec(0, bold=T) %>% add_footnote(c("Source: Author's elaboration", "TC: Times Cited"), notation ="number") 
```

# 6. Performance Analysis

## 6.4. Journals

```{r Influencial SO, include=T}

influ_SO <- read_csv(here("Data", "Processed", "influ_SO.csv"), show_col_types = F)

kbl(influ_SO, "latex", caption = "Journals' Performance") %>% 
  # column_spec(4, latex_column_spec = "p{5cm}") %>% 
kable_styling(bootstrap_options = "striped", full_width = F, font_size = 8) %>%
  row_spec(0, bold=T) %>%
  add_footnote(c("Source: Author's elaboration", "TC: Times Cited", "PD: Published Documents"), notation ="number")


```

# 6. Performance Analysis

## 6.5. Affiliations/ Universities

```{r Influencial UN, include=T, results='asis', echo=F}


influ_UN <- read_csv(here("Data", "Processed", "univ_perf.csv"), show_col_types = F)


kbl(influ_UN, "latex", caption = "Top 10 University's Performance") %>%
  # column_spec(4, latex_column_spec = "p{5cm}") %>%
kable_styling(bootstrap_options = "striped", full_width = F, font_size = 8) %>%
  row_spec(0, bold=T) %>%
  add_footnote(c("Source: Author's elaboration", "TC: Times Cited", "PD: Published Documents"), notation ="number")


```

# 6. Performance Analysis

## 6.5. Countrys' Performance

```{r Influencial CO, include=T}
grid.arrange(gg1,gg.mid,gg2,ncol=3,widths=c(4/9,1/6,4/9))


```

# 7. Science Mapping

## Similarity measures

Following @kammerer2021

-   **Research base:** cluster of academic publications in a research
    field that are considered fundamental to the development and
    understanding of the field.

-   **Research front:** cluster of academic publications that refers to
    emerging active areas of research considering themselves with a
    similar unsolved research problem.

```{=tex}
\begin{center}
\includegraphics[width=0.6\textwidth]{pic_1.png}
\end{center}
```
# 7. Science Mapping

\vspace{0.5cm}

This network shows the emergence of the literature on the digital divide

```{r Co_cite_P1,include=T}

load(here("Script", "Environments",  "Matrix_4_Net_env.RData"))

set.seed(30100)
net=netPlot(CC1_mat, n = 35, Title = "Co-Citation Network,  2000 - 2007", # Louvain
                type = "auto", size.cex=TRUE, size=20, 
                remove.multiple=FALSE, labelsize=0.9,edgesize = 4, edges.min=1, label = T,
            cluster = "optimal", curved = F )

```

# 7. Science Mapping

```{r Co_cite_P2,include=T}

load(here("Script", "Environments",  "Matrix_4_Net_env.RData"))


set.seed(1030034)
net=netPlot(CC2_mat, n = 25, Title = "Co-Citation Network, 2008 - 2015", 
            type = "auto", size.cex=TRUE, size=20, 
            remove.multiple=FALSE, labelsize=0.9,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain")

```

# 7. Science Mapping

```{r Co_cite_P3,include=T}

set.seed(1003)
net=netPlot(CC3_mat, n = 30, Title = "Co-Citation Network P3 louvain", 
            type = "auto", size.cex=TRUE, size=20, label.color = F,
            remove.multiple=FALSE, labelsize=0.7,edgesize = 4, edges.min=1, label = T,
            cluster = "louvain")

```

# References {.allowframebreaks}
