---
title: "Test table"
author: "Luis Carlos Castillo"
date: "`r Sys.Date()`"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F )
```

```{r Loading packages, include=FALSE}
library(tidyverse)
library(devtools)
library(here)
library(bibliometrix)
library(stringr)
library(readr)
library(visdat)
library(stringr)
library(dplyr)
library(readxl)
library(janitor)
library(ggplot2)
library(gridExtra)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(rmarkdown)
library(tidytext)
library(knitr)
library(qwraps2)
library(dplyr)
library(tables)
library(bibtex)
library(igraph)
library(RColorBrewer)
library(Matrix)
library(flextable)
library(ggpubr)
library(cowplot)
library(ggeasy)
library(Hmisc)
library(gridGraphics)
library(gridBase)

```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r Descriptive, echo=FALSE, include=T}
desc_a <- read_csv(here("Data", "Processed", "desc_analysis.csv"))


# set_flextable_defaults(
#   border.color = "gray")
# 
# flextable(head(desc_a)) %>% 
#   bold(part = "header") %>% 
#   add_footer_lines("Descriptive Analysis") %>% 
#   theme_vanilla()

kable(desc_a,"latex") %>% 
  kable_styling(latex_options="scale_down") %>% 
  row_spec(c(0), bold=T) %>% row_spec(c(5), color = "blue") #005997
  
   

```
